Transition path sampling based calculation of free energies of enzymatic reactions

### 2022-04-08 tex/manuscript.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) has been widely applied for the
  calculations of transition state and reaction coordinates of 
  condensed phase systems. Here, we discuss a scheme for the calculation
  of free energies with the TPS method 
  We discuss a scheme for the calculation of free energies within 
  the transition path sampling (TPS) method. TPS is often used for
  the calculations of reaction coordinates of complex reactions in 
  the condensed phase and in this work we use it to sample the 
  equilibrium distribution of the reaction coordinate. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the human Mat2a enzyme and the plasmodium 
  adenosine deaminase to demonstrate the applicability of this method
  for enzyme catalysis reactions.

  %This is an example document for the \textsf{achemso} document
  %class, intended for submissions to the American Chemical Society
  %for publication. The class is based on the standard \LaTeXe\
  %\textsf{report} file, and does not seek to reproduce the appearance
  %of a published paper.

  %This is an abstract for the \textsf{achemso} document class
  %demonstration document.  An abstract is only allowed for certain
  %manuscript types.  The selection of \texttt{journal} and
  %\texttt{manuscript} will determine if an abstract is valid.  If
  %not, the class will issue an appropriate error.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}

@@

CHARMM force field \cite{Brooks09JComputChem30p1545}

%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
The SN2 reaction catalyzed by the human Mat2a enzyme leads to the production of 
S-Adenosylmethionine compound which is part of the ATP cycle \cite{Niland21Biochem60p791}
\begin{figure}[ht!]
\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=0.5]{figures/comm-80-mat2a.pdf}
\end{figure}

@@

%\begin{figure}[ht]
%\centering
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/reactant-state.png}
%\caption{Reactant state}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/int1.png}
%\caption{Intermediate state}
%\label{fig:minipage2}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/product-state.png}
%\caption{Product state}
%\label{fig:minipage1}
%\end{minipage}
%\end{figure}

%\subsection{Outline}
%\begin{figure}[ht!]
%\centering
%\begin{minipage}[b]{0.45\linewidth}
%\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
%\caption{Trajectory 167}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.45\linewidth}
%\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
%\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
%\label{fig:minipage2}
%\end{minipage}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) has been widely applied for the
  calculations of transition state and reaction coordinates of 
  condensed phase systems. Here, we discuss a scheme for the calculation
  of free energies with the TPS method 
  We discuss a scheme for the calculation of free energies within 
  the transition path sampling (TPS) method. TPS is often used for
  the calculations of reaction coordinates of complex reactions in 
  the condensed phase and in this work we use it to sample the 
  equilibrium distribution of the reaction coordinate. This 
  distribution can then be used for the calculations of free energies
  using standard Boltzmann inversion schemes. We calculate the free
  energy profiles of the reactions catalyzed by the human methionine 
  S-adenosyltransferase (MAT2A) enzyme and the plasmodium 
  vivax adenosine deaminase enzyme to demonstrate the applicability of this 
  method. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}

@@

%------------------------------------------------------------------------------ We performed computational studies on two enzymes where we have shown
that fast dynamics is part of the reaction mechanism and calculated the electric field near
the bond-breaking event.
QMMM simulations were used for the TPS simulations in this work and the 
CHARMM force field \cite{Brooks09JComputChem30p1545} is used for the 
MM portion and the PM3 semiempirical method \cite{Repasky02JComputChem23p1601} 
is used for the QM region.

%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
reaction between Adenosine triphosphaste and L-methionine which leads to the 
formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
The mechanism of this reaction is determined to follow a SN$_2$ transition state
with the sulphur atom of the L-methionine as the nucleophile with the 
triphosphate leaving group.

\begin{figure}[ht!]
\includegraphics[scale=0.25]{figures/mat2a-transition-state.png}
\end{figure}

\begin{figure}[ht!]
\includegraphics[scale=0.5]{figures/comm-60-mat2a.pdf}
\end{figure}

@@

%\begin{figure}[ht]
%\centering
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/reactant-state.png}
%\caption{Reactant state}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/int1.png}
%\caption{Intermediate state}
%\label{fig:minipage2}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.32\linewidth}
%\includegraphics[scale=0.3]{figures/product-state.png}
%\caption{Product state}
%\label{fig:minipage1}
%\end{minipage}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
